---
title: "project4_regtree"
author: "Akhil Havaldar"
date: "11/9/2022"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## a) 
- Need to change the quality column to a factor.
- Need to rename columns to remove spaces from the variable names

## b) 
```{r}
library(readr)
library(tree)
library(janitor)

wine <- read_csv("wine.csv")
wine$quality <- factor(wine$quality)
wine <- clean_names(wine)       # removes spaces from names

set.seed(4630)
sample.data<-sample.int(nrow(wine), floor(0.5*nrow(wine)), replace = F)
train<-wine[sample.data, ]
test<-wine[-sample.data, ]

y.test<-test[,"alcohol"]
```

## b - output of summary function
```{r}
tree1 <- tree(alcohol~. , train)
summary(tree1)
```
## b - terminal nodes
- there are 13 terminal nodes

## b - predictors used
- density, fixed acidity, residual sugar, suplhates, quality, and citric acid

## b - graph
```{r}
plot(tree1)
text(tree1, cex=0.6) 
```
## b - test mse
```{r}
tree1.pred<-predict(tree1, newdata =test)
mse.tree1 <- mean((tree1.pred-y.test$alcohol)^2)
print(mse.tree1) 
```
## c - Pruned tree
```{r}
set.seed(4630)
cv.wine<-cv.tree(tree1, K=10)
cv.wine
plot(cv.wine$size, cv.wine$dev,type='b')

trees.num.wine<-cv.wine$size[which.min(cv.wine$dev)]
trees.num.wine

tree.full<-tree::tree(alcohol~., data = train)
prune.full<-tree::prune.tree(tree.full, best=trees.num.wine)
summary(prune.full)
```
## c - graph
```{r}
plot(prune.full)
text(prune.full, cex=0.5)
```

## c - test mse
```{r}
tree.prune.pred<-predict(prune.full, newdata =test)
mse.tree2 <- mean((tree.prune.pred-y.test$alcohol)^2)
print(mse.tree1) 
```

## d - random forest
```{r}
library(randomForest) 
rf<-randomForest::randomForest(alcohol~., data=train, mtry=3,importance=TRUE)
rf

importance(rf)
varImpPlot(rf)
```

## d- test mse
```{r}
rf.pred<-predict(rf, newdata =test)
mse.rf <- mean((rf.pred-y.test$alcohol)^2)
print(mse.rf) 
```

## e - table
```{r}
out <- data.frame(mse.tree1, mse.tree2, mse.rf)
colnames(out) <- c("MSE.RecBinary", "MSE.Pruned", "MSE.RandomForest")
out
```

- Main conclusions found in PDF 

